\vspace{-2.5em} Singular value decomposition of single view\\ polarized fluorescence microscopes\vspace{-1em}

### 2018-05-23 notes/2018-05-14-single-view-continuous-svd/report/report.tex
Patrick and Corey's corrections from group meeting.

\subsection{Singular value decomposition}
Plugging the transfer function into the frequency-domain eigenvalue problem (Eq.
\ref{eq:freqsvd}) yields a $3\times 3$ eigenvalue problem (we suppress the index
$\bs{\rho}$ and understand that we need to solve this eigenvalue problem at each
value of $\bs{\rho}$)
\begin{align}
  \mathsf{K}\mathsf{V}_j &= \mu_{j}\mathsf{V}_j,\\
  \begin{bmatrix}
    \sum_{l,m}H_{l,-2}^{m}H_{l,-2}^{m}&\sum_{l,m}H_{l,-2}^{m}H_{l,0}^{m}&\sum_{l,m}H_{l,-2}^{m}H_{l,2}^{m}\\
    \sum_{l,m}H_{l,0}^{m}H_{l,-2}^{m}&\sum_{l,m}H_{l,0}^{m}H_{l,0}^{m}&\sum_{l,m}H_{l,0}^{m}H_{l,2}^{m}\\
    \sum_{l,m}H_{l,2}^{m}H_{l,-2}^{m}&\sum_{l,m}H_{l,2}^{m}H_{l,0}^{m}&\sum_{l,m}H_{l,2}^{m}H_{l,2}^{m}\\    
  \end{bmatrix}\mathsf{V}_j
&= \mu_{j}\mathsf{V}_j,\\
  \begin{bmatrix}
    \{H_{2,-2}^{-2}\}^2 + \{H_{4,-2}^{-2}\}^2&0&0\\
    0&\{H_{0,0}^{0}\}^2 + \{H_{2,0}^{0}\}^2 + \{H_{4,0}^{0}\}^2&0\\
    0&0&\{H_{2,2}^{2}\}^2 + \{H_{4,2}^{2}\}^2\\    
  \end{bmatrix}\mathsf{V}_j
&= \mu_{j}\mathsf{V}_j.
\end{align}
A diagonal matrix has its eigenvalues along the diagonal and its eigenvectors as
unit vectors so
\begin{align}
  \mu_{\bs{\rho},0} &= \{H_{0,0}^{0}(\rho)\}^2 + \{H_{2,0}^{0}(\rho)\}^2 + \{H_{4,0}^{0}(\rho)\}^2,\\
  \mu_{\bs{\rho},1} = \mu_{\bs{\rho},2} &= \{H_{2,2}^{2}(\rho)\}^2 + \{H_{4,2}^{2}(\rho)\}^2,
\end{align}
and
\begin{align}
  \mathsf{V}_0 =
  \begin{bmatrix}
    0\\1\\0\\
  \end{bmatrix},\qquad
  \mathsf{V}_1 =
  \begin{bmatrix}
    1\\0\\0\\
  \end{bmatrix},\qquad
  \mathsf{V}_2 =
  \begin{bmatrix}
    0\\0\\1\\
  \end{bmatrix}.
\end{align}
Notice that we have two degenerate eigenvalues, but we have found trivial
orthonormal eigenvectors---in general we will need to apply the Gram-Schmidt
procedure.

@@

\appendix
\section{Calculating products of circular and spherical harmonics}
To calculate the kernels of arbitrary microscope designs we need an efficient
way to calculate the products of functions that are linear combinations of
circular and spherical harmonics. We start by considering the simplest case of
multiplying two linear combinations of circular harmonics given by
\begin{align}
  f(\mh{p}) = \sum_{n=0}^\infty c_nz_n(\mh{p}), \qquad f'(\mh{p}) = \sum_{n'=0}^\infty c_{n'}'z_{n'}(\mh{p}). 
\end{align}
The product of these two functions is given by 
\begin{align}
  f''(\mh{p}) &= f(\mh{p})f'(\mh{p})=\sum_{n=0}^\infty \sum_{n'=0}^\infty c_n c_{n'}' z_n(\mh{p}) z_{n'}(\mh{p}) = \sum_{n''=0}^{\infty} c_{n''}''z_{n''}(\mh{p}), \label{eq:two}
  \end{align}
  and we would like to find a relationship between the input coefficients ($c_n$
  $c_{n'}'$) and the output coefficients $c_{n''}''$. To find this relationship
  we can rewrite the product the circular harmonics as a linear combination of
  circular harmonics
  \begin{align}
    z_n(\mh{p})z_{n'}(\mh{p}) = \sum_{j''=0}^{\infty}P_{n,n',n''}z_{j''}(\mh{p})\label{eq:product}
  \end{align}
  where $P^{n''}_{n,n'}$ are the triple integrals of the circular harmonics
  \begin{align}
    P_{n,n',n''} = \int_{\mbb{S}^1}d\mh{p}\, z_{n}(\mh{p})z_{n'}(\mh{p})z_{n''}(\mh{p}).\label{eq:triple}
  \end{align}
  Plugging Eq. \ref{eq:product} into Eq. \ref{eq:two} gives
  \begin{align}
    \sum_{n=0}^\infty \sum_{n'=0}^\infty c_n c_{n'}' \left[\sum_{j''=0}^{\infty}P^{n''}_{n,n'}z_{j''}(\mh{p})\right] = \sum_{n''=0}^{\infty} c_{n''}''z_{n''}(\mh{p}).    
  \end{align}
  Therefore, the coefficients are related by
  \begin{align}
    c_{n''}'' = \sum_{n=0}^\infty \sum_{n'=0}^\infty P_{n,n',n''} c_n c_{n'}'.
  \end{align}
  Rewriting this equation in Einstein notation (summation over matching indices
  is implied, upper indices are ``column'' indices, lower indices are ``row''
  indices) gives
  \begin{align}
    {c''}^{n''} = P^{n''}_{n,n'} c_n c_{n'}'.
  \end{align}
  We can precompute the triple integrals $P_{n,n',n''}$ using a symbolic package
  like Sympy.

@@

  To calculate the kernel for general polarized light microscopes we will need to multiply functions in the following form
  \begin{align}
  f(\mh{p}, \mh{s}) = \sum_{n=0}^\infty\sum_{j=0}^\infty c_{n,j} z_n(\mh{p})y_j(\mh{s}), \qquad f'(\mh{p}, \mh{s}) = \sum_{n'=0}^\infty\sum_{j'=0}^\infty c_{n',j'}' z_n(\mh{p})y_j(\mh{s}).
\end{align}
The product will be in the form
\begin{align}
  f(\mh{p}, \mh{s})f'(\mh{p}, \mh{s}) = \sum_{n=0}^\infty\sum_{j=0}^\infty c_{n'',j''}'' z_{n''}(\mh{p})y_{j''}(\mh{s}),
\end{align}
where
\begin{align}
  {c''}^{n'', j''} = P_{n,n'}^{n''} G_{j,j'}^{j''} c_{n,j} c_{n', j'}'.\label{eq:final}
\end{align}
Equation \ref{eq:final} is the main result of this section. It shows that we can
precalculate the triple integrals of the circular and spherical harmonics and
use the results to efficiently find the coefficients of the product of two
arbitrary kernels. We can think of Eq. \ref{eq:final} as a bilinear map that
acts within the vector space of harmonic function coefficients. The bilinear map
takes two elements of the vector space and maps them to another element of the
vector space by a rank-6 tensor product. A lower dimensional example of a
bilinear map is the cross product which takes two vectors in three-dimensional
Euclidean space and maps them to another vector by a rank-3 tensor product.
\end{document}

>>>>

\subsection{Singular value decomposition}
Plugging the transfer function into the frequency-domain eigenvalue problem (Eq.
\ref{eq:freqsvd}) yields a denumerably infinite eigenvalue problem with only
$3\times 3$ non-zero entries in the matrix (we suppress the index $\bs{\rho}$
and understand that we need to solve this eigenvalue problem at each value of
$\bs{\rho}$)
\begin{align}
  \mathsf{K}\mathsf{V}_j &= \mu_{j}\mathsf{V}_j,\\
  \begin{bmatrix}
    \sum_{l,m}H_{l,-2}^{m}H_{l,-2}^{m}&\sum_{l,m}H_{l,-2}^{m}H_{l,0}^{m}&\sum_{l,m}H_{l,-2}^{m}H_{l,2}^{m}\\
    \sum_{l,m}H_{l,0}^{m}H_{l,-2}^{m}&\sum_{l,m}H_{l,0}^{m}H_{l,0}^{m}&\sum_{l,m}H_{l,0}^{m}H_{l,2}^{m}\\
    \sum_{l,m}H_{l,2}^{m}H_{l,-2}^{m}&\sum_{l,m}H_{l,2}^{m}H_{l,0}^{m}&\sum_{l,m}H_{l,2}^{m}H_{l,2}^{m}\\    
  \end{bmatrix}\mathsf{V}_j
&= \mu_{j}\mathsf{V}_j,\\
  \begin{bmatrix}
    \{H_{2,-2}^{-2}\}^2 + \{H_{4,-2}^{-2}\}^2&0&0\\
    0&\{H_{0,0}^{0}\}^2 + \{H_{2,0}^{0}\}^2 + \{H_{4,0}^{0}\}^2&0\\
    0&0&\{H_{2,2}^{2}\}^2 + \{H_{4,2}^{2}\}^2\\    
  \end{bmatrix}\mathsf{V}_j
&= \mu_{j}\mathsf{V}_j.
\end{align}
A diagonal matrix has its eigenvalues along the diagonal and its eigenvectors as
unit vectors so
\begin{align}
  \mu_{\bs{\rho},0} &= \{H_{0,0}^{0}(\rho)\}^2 + \{H_{2,0}^{0}(\rho)\}^2 + \{H_{4,0}^{0}(\rho)\}^2,\\
  \mu_{\bs{\rho},1} = \mu_{\bs{\rho},2} &= \{H_{2,2}^{2}(\rho)\}^2 + \{H_{4,2}^{2}(\rho)\}^2,
\end{align}
and
\begin{align}
  \mathsf{V}_0 =
  \begin{bmatrix}
    0\\1\\0\\
  \end{bmatrix},\qquad
  \mathsf{V}_1 =
  \begin{bmatrix}
    1\\0\\0\\
  \end{bmatrix},\qquad
  \mathsf{V}_2 =
  \begin{bmatrix}
    0\\0\\1\\
  \end{bmatrix}.
\end{align}
Notice that we have two degenerate eigenvalues, but we have found trivial
orthonormal eigenvectors---in general we will need to apply the Gram-Schmidt
procedure.

@@

\appendix
\section{Calculating products of circular and spherical harmonics}
To calculate the kernels of arbitrary microscope designs we need an efficient
way to calculate the products of functions that are linear combinations of
circular and spherical harmonics. We start by considering the simplest case of
multiplying two linear combinations of circular harmonics given by
\begin{align}
  f(\mh{p}) = \sum_{n=0}^\infty c_nz_n(\mh{p}), \qquad f'(\mh{p}) = \sum_{n'=0}^\infty c_{n'}'z_{n'}(\mh{p}). 
\end{align}
The product of these two functions is given by 
\begin{align}
  f''(\mh{p}) &= f(\mh{p})f'(\mh{p})=\sum_{n=0}^\infty \sum_{n'=0}^\infty c_n c_{n'}' z_n(\mh{p}) z_{n'}(\mh{p}) = \sum_{n''=0}^{\infty} c_{n''}''z_{n''}(\mh{p}), \label{eq:two}
  \end{align}
  and we would like to find a relationship between the input coefficients ($c_n$
  $c_{n'}'$) and the output coefficients $c_{n''}''$. To find this relationship
  we can rewrite the product the circular harmonics as a linear combination of
  circular harmonics
  \begin{align}
    z_n(\mh{p})z_{n'}(\mh{p}) = \sum_{j''=0}^{\infty}P_{n,n',n''}z_{j''}(\mh{p})\label{eq:product}
  \end{align}
  where $P^{n''}_{n,n'}$ are the triple integrals of the circular harmonics
  \begin{align}
    P_{n,n',n''} = \int_{\mbb{S}^1}d\mh{p}\, z_{n}(\mh{p})z_{n'}(\mh{p})z_{n''}(\mh{p}).\label{eq:triple}
  \end{align}
  Plugging Eq. \ref{eq:product} into Eq. \ref{eq:two} gives
  \begin{align}
    \sum_{n=0}^\infty \sum_{n'=0}^\infty c_n c_{n'}' \left[\sum_{j''=0}^{\infty}P^{n''}_{n,n'}z_{j''}(\mh{p})\right] = \sum_{n''=0}^{\infty} c_{n''}''z_{n''}(\mh{p}).    
  \end{align}
  Therefore, the coefficients are related by
  \begin{align}
    c_{n''}'' = \sum_{n=0}^\infty \sum_{n'=0}^\infty P_{n,n',n''} c_n c_{n'}'.
  \end{align}
  Rewriting this equation in Einstein notation (summation over matching indices
  is implied, upper indices are ``column'' indices, lower indices are ``row''
  indices) gives
  \begin{align}
    {c''}^{n''} = P^{n''}_{n,n'} c^n {c'}^{n'}.
  \end{align}
  We can precompute the triple integrals $P_{n,n',n''}$ using a symbolic package
  like Sympy.

@@

  To calculate the kernel for general polarized light microscopes we will need to multiply functions in the following form
  \begin{align}
  f(\mh{p}, \mh{s}) = \sum_{n=0}^\infty\sum_{j=0}^\infty c_{n,j} z_n(\mh{p})y_j(\mh{s}), \qquad f'(\mh{p}, \mh{s}) = \sum_{n'=0}^\infty\sum_{j'=0}^\infty c_{n',j'}' z_n(\mh{p})y_j(\mh{s}).
\end{align}
The product will be in the form
\begin{align}
  f(\mh{p}, \mh{s})f'(\mh{p}, \mh{s}) = \sum_{n=0}^\infty\sum_{j=0}^\infty c_{n'',j''}'' z_{n''}(\mh{p})y_{j''}(\mh{s}),
\end{align}
where
\begin{align}
  {c''}^{n'', j''} = P_{n,n'}^{n''} G_{j,j'}^{j''} c^{n,j} {c'}^{n', j'}.\label{eq:final}
\end{align}
Equation \ref{eq:final} is the main result of this section. It shows that we can
precalculate the triple integrals of the circular and spherical harmonics and
use the results to efficiently find the coefficients of the product of two
arbitrary kernels. We can think of Eq. \ref{eq:final} as a bilinear map that
acts within the vector space of harmonic function coefficients. The bilinear map
takes two elements of the vector space and maps them to another element of the
vector space by a rank-6 tensor product. A lower dimensional example of a
bilinear map is the cross product which takes two vectors in three-dimensional
Euclidean space and maps them to another vector by a rank-3 tensor product.
\end{document}
